Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-16 paper/sheaf_lics.tex
pass on section 5.1

\newcommand{\IsType}[1]{\mathop{\mathrm{Is\mymathhyphen}#1\mathrm{\mymathhyphen
      type}} }

\newcommand{\modal}{\ensuremath{\ocircle}}
\newcommand \True {\mathrm{true}}
\newcommand \False {\mathrm{false}}
\newcommand \closure[1] {\overline{#1}}
\newcommand \Char[1] {\chi_{#1}}%{\mathrm{char}(#1)}
\newcommand \E {\mathcal{E}}
\newcommand \Hom[1] {\mathrm{Hom}_{#1}}
\newcommand \Sh[1] {\mathrm{Sh}_{#1}}
\newcommand \squash[1] {\| #1 \| }
\newcommand \fib[2] {\mathrm{fib}_{#1}(#2)}
\newcommand \colim[1] {\mathrm{colim}(Q)}
\newcommand \zero {\mathbf{0}}
\newcommand \one {\mathbf{1}}

@@

The central result used here is the following
\begin{lem}
  Let $X:\Type_p$, $Y:\Type_p\modal$ and $f:X\to Y$. If for all $y:Y$,
  $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
\end{lem}
%
Here, $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, thus we
can define $$\phi:\modal \left(\sum_{x:X}  (f x = y)\right) \to
\sum_{x:\modal X} (\modal f x = \eta_Y y).$$
We just need to check that the image of its fibers by $\modal$ are
contractible.
%
From definition of left exactness, one can deduce the following~:
\begin{lem}
  Let $A,B:\Type_p$. Let $f:A\to B$. If $\modal A$ and $\modal B$ are
  contractible, then so is $\fib f b$ for any $b:B$.
\end{lem}
We can rewrite $\fib \phi {x;p}$ in terms of the fiber of a fonction
$\fib {\eta_X} a \to \fib {\eta_Y} b$.

@@

Finally, $\modal(\fib \phi {x;p})$ is contractible, and the result is proved.
\kq{Finish that}
\end{proof}

@@

\begin{defi}
  For any type $A$ and $E$, $\iota : A -> E$ and $F:\Type_{n+1}$, we define
    $$
    \Phi_E^\iota : (E \to F) \to (A \to F) 
    $$
    %
    as the map sending an
    arrow $f:E\to F$ to its restriction the restriction of $f \circ m$.
  % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
  %   map sending $f~:~E \to f$ to $\restriction f A$.
  % \end{itemize}
\end{defi}

@@

\nt{Il faut mieux introduire ces propriétés, $\Type_n^\modal$ is a sheaf
  est très importante en particulier. Il faut qu'on comprenne d'où ça
  vient et que tu fasses une preuve, même un peu idalisée. Je dirais
  que tu peux mettre des lemmes genres nj\_inter en disant que telle
  flèche est un iso, and co.
  En tout cas, cette partie peut et doit être bien plus longue.
}

\begin{prop}\label{prop:sheaf-forall}
  If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
  $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
\end{prop}

\begin{prop}\label{prop:sheaf-is-sheaf}
  $\Type_n^\modal$ is a sheaf.
\end{prop}
\begin{proof}
  Firstly, we prove that $\Type_n^\modal$ is separated. Let $E:\Type$
  and $\chi:E \to \Type_n$, dense in $E$. Let $\phi_1,\phi_2:E \to
  \Type_n^\modal$, such that $\phi_1 \circ \pi_1 = \phi_2 \circ
  \pi_1$. Finally, let $x:E$. We show $\phi_1 x = \phi_2 x$ using
  univalence.

  As $\chi$ is dense, we have a term $m_x : \modal(\chi\, x)$.
  But as $\phi_2 x$ is modal, thus we can obtain a term $h_x : \chi\,
  x$. 
  As $\phi_1$ and $\phi_2$ are equal on $\sum_{e:E}\chi\, e$, we
  have an arrow $\phi_1 x \to \phi_2 x$.
  The same method leads to an arrow $\phi_2 x \to \phi_1 x$, and we
  prove they are each other inverse.

@@

@@

From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
induction the truncation index $n$, a
left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
sense that the following property holds.
%
% We have to be careful~: if we want to propagate the properties of the
% first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
% with $\modal_{-1}$, \ie we want the property
\begin{prop}\label{prop:hprop}
  For any mere proposition $P$ (with $\widehat P$ is $P$ seen as a
  $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
  following coherence diagram commutes
  $$\xymatrix{
    P \ar@{=}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
    \modal_{-1} P \ar@{=}[r] & \modal_n \widehat P 
  }$$
\end{prop}

Note that once the first modality is defined, the extension to all
$\Type_n$ is automatic. This means that the new logical principles
that can be added by the sheaf construction must be added in the
first modality already. The sheafification is just a way to propagate
those new principles to other types than just mere proposition.

@@

Let $T : \Type_{n+1}$. We define $\square T$ as the image (see
Section~\ref{sec:epi-mono-fact}) of
$\modal^T \circ \{\cdot\}_T$, as in
$$\xymatrix{
  T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal^T} \\
  \square T \ar[r]& \left( \Type_n^\modal \right)^T
}, $$
where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
(t':T),~t=t'$.
\nt{parler du système de factorization}
In type theory words, 
\begin{align*}
\square T &= \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')) \\
          &= \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:X} 
            (\lambda t,~\modal (a=t)) = u\right\|
\end{align*}

There again, the separation step has the same definition as
in~\cite{maclanemoerdijk}, using $\Type_n^\modal$ instead of the
$j$-subobject classifier.

\begin{prop}
  For any $T:\Type_{n+1}$, $\square T$ is separated.  
\end{prop}

\begin{proof}
We will use the following lemma~:
\begin{lem}
  A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
  into a separated $(n+1)$-truncated type $U$ is itself separated.
\end{lem}
As $\square T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
have to show that the latter is separated. But it is the case because
$\Type_n^\modal$ is a sheaf (by Proposition~\ref{prop:sheaf-is-sheaf})
and a function type is a sheaf as soon
as its codomain is a sheaf (by Proposition~\ref{prop:sheaf-forall}).
\end{proof}

We now need to show that $\square$ is universal. The following
lemma is central in the proof~:
\begin{lem}\label{lem:sepiscolim}
  Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
  diagonal diagram
  $$\xymatrix @C=4em  { 
    \cdots \overline{\Delta_3} \tar[r]{4pt} & \overline{\Delta_2} \dar[r]{2pt} &
    \overline{\Delta_1}
  }$$
where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
\end{lem}

This lemma is an adaptation of the sheafification process
in~\cite{maclanemoerdijk}, where they consider only the kernel pair of
$\mu_T$ instead of the \v{C}ech nerve.

\begin{proof}
  As $\mu_T$ is an surjection, using the particular case of the last
  Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
  we know that $\square T$ is the colimit
  of $C(\mu_T)$. If we can show that $C(\mu_T) = C(\id)$, the the
  result will follow.

  Let $k:\nat$, let's show that 
  \begin{align*}
    &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
      \land \mu_T t_{k-1} \mu_T t_k)\\
    &= \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
      \land t_{k-1} = t_k)
  \end{align*}
By induction on $k$, and the preservation of products by $\modal$, it
suffices to show that for any $a,b:T$, $\modal (a=b) = (\mu_T a =
\mu_T b)$. By univalence, we want arrows in both ways, forming an
equivalence.
\begin{itemize}
\item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
  first components yields $q : \prod_{t:T} \modal(a=t) = \modal (b=t)
  $.
  Taking for example $t=b$, we deduce $\modal (a=b) = \modal(a=a)$,
  and the latter is inhabited by $\eta_{a=a} 1$.
\item Suppose now $p : \modal(a=b)$. Let $\iota$ be the first
  projection from $\square T \to (T \to \Type_n^\modal)$. $\iota$ is
  an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
  (\mu_T b)$, \ie $\prod_{t:T}\modal (a=t) = \modal (b=t)$. The latter
  remains true by univalence.
\end{itemize}
The fact that these two form an equivalence is technical, we refer to
the formalization for an explicit proof.
\end{proof}

Now, let $Q$ be a separated $\Type_{n+1}$, and $f:T \to Q$. Then the
following diagram commutes
$$\xymatrix @C=4em{ 
    \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
    \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
    \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
    & Q &
  }$$
But we know (lemma~\ref{lem:sepiscolim}) that $\square T$ is the
colimit of the closed diagonals diagram, thus there is an universal
arrow $\square T \to Q$.

This proves the following proposition
\begin{prop}
  $(\square,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
\end{prop}

We now have to prove point {\it (v)} in the definition of
modality. Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be
a sheaf family. We want to show that $\sum_{x:A} (Bx)$ is
separated. Let $E$ be a type, and $\sum_{e:E} (\chi\,e)$ a dense
subobject of E.

@@

$$\xymatrix@C=8em@R=4em{
  \sum_{y:\sum_{e:E} (\chi\,
  e)} x=y_1 \ar@<2pt>[r]^{\qquad \pi_2\circ f\circ\pi_1\circ \pi_1}
\ar@<-2pt>[r]_{\qquad \pi_2\circ g\circ \pi_1\circ \pi_1}
\ar[d]_{\mathrm{dense}}& B\,x \\
  \sum_{y:E} x = y \ar@<2pt>[ur]^{\pi_2\circ f\circ \pi_1} \ar@<-2pt>[ur]_{\pi_2\circ g\circ \pi_1}&
}$$
Using the sepatation property of $B\,x$, one can show that second
components, transported correctly along the first components equality,
are equal. The complete proof can be found in the formalisation.
This proves the following proposition
\begin{prop}
  $(\square,\mu)$ defines a modality on $\Type_{n+1}$.
\end{prop}

As this modality is just a step in the construction, we will not show
that it is left exact, we will do it only for the sheafification modality.

@@

As any separated type $T$ embeds in $\left(\Type_n^\modal\right)^T$,
it suffices to take the closure of $T$ to get a sheaf, and the unit
$\nu_T:T \to \star T$ is obvious.
\begin{prop}
  $(\star,\nu)$ defines a reflective subuniverse.
\end{prop}
\begin{proof}
  Let $T,Q:\Type_{n+1}$ such that $Q$ is a sheaf. Let $f:T\to Q$.
  Because $Q$ is a sheaf, it is in particular separated ; thus we can
  extend $f$ to $\square f:\square T\to Q$.

  But as $\star T$ is the closure of $\square T$, $\square T$ is dense
  into $\star T$, so the sheaf property of $Q$ allows to extend
  $\square f$ to $\star f:\star T \to Q$.

@@

At total, with other files containing technical lemmas, the project
contains 6322 lines.
%\nt{describe how many line for each, be clear about what has not been formalized}

In our formalization, some things are admitted, either because they
are just an adaptation of a proposition already formalized
in~\cite{hottbook}, either because they are true, but very technical~:
\begin{itemize}
\item in \texttt{modalities.v}, an equality is admitted in the proof
  of proposition~\ref{sec:defin-basic-prop}.
\item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
  m)$ is admitted.
\item in \texttt{sheaf\_induction.v}, in the proof that
  $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
  of equality of projections.
\item in \texttt{sheaf\_induction.v}, we admit some very technical
  proofs in proofs that $\square$ and $\star$ are modalities.
\item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
  because of universes issues.
\end{itemize}

Some other lemmas are proved, then admitted. We do it so that Coq
doesn't constrain too much universes, without having to explicitly
annotate universes everywhere in the formalization.

The main issue in this formalization is the gestion of universes in
Coq~\cite{sozeau2014universe}. If Coq can handle cumulativity on all
$\Type$, it is not the case for the sigma type $\Type_n$. This issue
can be solved by adding an axiom of cumulativity for $\Type_n$, but as
it would not have any computational content, it would really
complicates the proofs.

Moreover, even the definition of successive sheafification types is
not allowed~:

\[ \left\{
    \begin{array}{l}
      \modal_{-1} = \lambda T:\HProp,~ \lnot\lnot T \\
      \displaystyle{\modal_{n+1} = \lambda T:\Type_{n+1}},\\
      \quad \displaystyle{\sum_{u:T \to \Type_n^\modal} \modal_{-1} 
      \left\|
      \sum_{a:T} u = (\lambda t,~\modal_n (a=t))
      \right\|}
    \end{array}
  \right.,\]
because Coq asks for every $\modal_n$ to have the same universes levels.

>>>>

\newcommand{\IsType}[1]
{\mathop{\mathrm{Is\mymathhyphen}#1\mathrm{\mymathhyphen type}} }

\newcommand{\modal}{\ensuremath{\ocircle}}
\newcommand \True {\mathrm{true}}
\newcommand \False {\mathrm{false}}
\newcommand \closure[1] {\overline{#1}}
\newcommand \Char[1] {\chi_{#1}}%{\mathrm{char}(#1)}
\newcommand \E {\mathcal{E}}
\newcommand \Hom[1] {\mathrm{Hom}_{#1}}
\newcommand \Sh[1] {\mathrm{Sh}_{#1}}
\newcommand \squash[1] {\| #1 \| }
\newcommand \separated {\mathop{\square} }
\newcommand \fib[2] {\mathrm{fib}_{#1}(#2)}
\newcommand \colim[1] {\mathrm{colim}(Q)}
\newcommand \zero {\mathbf{0}}
\newcommand \one {\mathbf{1}}

@@

The proof is based on the following central result.
\begin{lem}
  Let $X:\Type_p$, $Y:\Type_p\modal$ and $f:X\to Y$. If for all $y:Y$,
  $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
\end{lem}
%
As $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, we
can define $$\phi:\modal \left(\sum_{x:X}  (f x = y)\right) \to
\sum_{x:\modal X} (\modal f x = \eta_Y y).$$
We just need to check that the image of its fibers by $\modal$ are
contractible.
%
From definition of left exactness, one can deduce the following~:
\begin{lem}
  Let $A,B:\Type_p$. Let $f:A\to B$. If $\modal A$ and $\modal B$ are
  contractible, then so is $\fib f b$ for any $b:B$.
\end{lem}
We can rewrite $\fib \phi {x;p}$ in terms of the fiber of a fonction
$\fib {\eta_X} a \to \fib {\eta_Y} b$.

@@

Finally, $\modal(\fib \phi {x;p})$ is contractible, and the result is proved.
\kq{Finish that}

\nt{Indeed, I can't follow the proof for the moment}
\end{proof}

@@

\begin{defi}
  For any type $A$ and $E$, $\iota : A \to E$ and $F:\Type_{n+1}$, we define
    $$
    \Phi_E^\iota : (E \to F) \to (A \to F) 
    $$
    %
    as the map sending an
    arrow $f:E\to F$ to its restriction the restriction of $f \circ m$.
  % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
  %   map sending $f~:~E \to f$ to $\restriction f A$.
  % \end{itemize}
\end{defi}

@@

A fundamental property in the construction of the sheafification is
the fact that the type of modal types at level $n$ defines a sheaf. 
%
This property initially holds in the topos theoretic construction, but
it requires much work here, as we have to deal with equalities. In
particular, this property uses in particular the specific definition
of separated types and sheaves, which shows that we had actually no
choice in their statement.

\begin{prop}\label{prop:sheaf-is-sheaf}
  $\Type_n^\modal$ is a sheaf.
\end{prop}
\begin{proof}
  First, we prove that $\Type_n^\modal$ is separated. Let $E:\Type$
  and $\chi:E \to \Type_n$, dense in $E$. Let $\phi_1,\phi_2:E \to
  \Type_n^\modal$, such that $\phi_1 \circ \pi_1 = \phi_2 \circ
  \pi_1$ and let $x:E$. We show $\phi_1 x = \phi_2 x$ using
  univalence.

  As $\chi$ is dense, we have a term $m_x : \modal(\chi\, x)$.
  But as $\phi_2 x$ is modal, we can obtain a term $h_x : \chi\,
  x$. 
  As $\phi_1$ and $\phi_2$ are equal on $\sum_{e:E}\chi\, e$, we
  have an arrow $\phi_1 x \to \phi_2 x$.
  The same method leads to an arrow $\phi_2 x \to \phi_1 x$, and we
  prove they are each other inverse.

@@

An other fundamental property on sheaves is that the type of (dependent)
functions is a sheaf as soon as its codomain is a sheaf.

\begin{prop}\label{prop:sheaf-forall}
  If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
  $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
\end{prop}
\begin{proof}
  \nt{add a short proof here}
\end{proof}

@@

From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
induction on the truncation index $n$, a
left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
sense that the following property holds.
%
% We have to be careful~: if we want to propagate the properties of the
% first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
% with $\modal_{-1}$, \ie we want the property
\begin{prop}\label{prop:hprop}
  For any mere proposition $P$ (with $\widehat P$ is $P$ seen as a
  $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
  following coherence diagram commutes
  $$\xymatrix{
    P \ar@{=}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
    \modal_{-1} P \ar@{=}[r] & \modal_n \widehat P 
  }$$
\end{prop}

Note that once the first modality $\modal_{-1}$ is defined, its
extension to all $\Type_n$ is automatic. This means that the new
logical principles that can be added by the sheaf construction must be
added in the first modality already. Sheafification is just a way
to propagate those new principles to types than are not mere
propositions.

@@

Let $T : \Type_{n+1}$. We define $\separated T$ as the image (see
Section~\ref{sec:epi-mono-fact}) of
$\modal^T \circ \{\cdot\}_T$, as in
$$\xymatrix{
  T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal^T} \\
  \separated T \ar[r]& \left( \Type_n^\modal \right)^T
}, $$
where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
(t':T),~t=t'$.
%
$\separated T$ can be given explicitly by
%
$$
\begin{array}{rcl}
\separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')) \\
          & \defeq & \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:X} 
            (\lambda t,~\modal (a=t)) = u\right\|.
\end{array}
$$
%
This corresponds to the free separated object used in the topos
theoretic construction, but using $\Type_n^\modal$ instead of the
$j$-subobject classifier $\Omega_j$.
%
\begin{prop}
  For any $T:\Type_{n+1}$, $\separated T$ is separated.  
\end{prop}

\begin{proof}
We use the following lemma:
\begin{lem}
  A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
  into a separated $(n+1)$-truncated type $U$ is itself separated.
\end{lem}
As $\separated T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
have to show that the latter is separated. But it is the case because
$\Type_n^\modal$ is a sheaf (by Proposition~\ref{prop:sheaf-is-sheaf})
and a function type is a sheaf as soon
as its codomain is a sheaf (by Proposition~\ref{prop:sheaf-forall}).
\end{proof}

To show that $\separated$ is a modality, we need t show that $\separated T$
is universal among separated type below $T$. This comes from the
following central lemma which connects $\separated T$ to the colimit of
the \v{C}ech nerve of $\mu_T$.

\begin{lem}\label{lem:sepiscolim}
  Let $T:\Type_{n+1}$. Then $\separated T$ is the colimit of the closed
  diagonal diagram
  $$\xymatrix @C=4em  { 
    \cdots \overline{\Delta_3} \tar[r]{4pt} & \overline{\Delta_2} \dar[r]{2pt} &
    \overline{\Delta_1}
  }$$
where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
\end{lem}

Again, this lemma is an adaptation of the sheafification process in
topos theory, where the kernel pair of $\mu_T$
is replaced by its \v{C}ech nerve.

\begin{proof}
  As $\mu_T$ is an surjection, using the particular case of 
  Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
  we know that $\separated T$ is the colimit of $C(\mu_T)$. 
%
  Thus the result follows from the fact that $C(\mu_T) = C(\id)$,
  which we now show.

  Let $k:\nat$, we will show that 
  \begin{align*}
    &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
      \land \mu_T t_{k-1} \mu_T t_k)\\
    &= \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
      \land t_{k-1} = t_k)
  \end{align*}
By induction on $k$, and the preservation of products by $\modal$, it
suffices to show that for any $a,b:T$, $\modal (a=b) = (\mu_T a =
\mu_T b)$. By univalence, we want arrows in both ways, forming an
equivalence.
\begin{itemize}
\item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
  first components yields $q : \prod_{t:T} \modal(a=t) = \modal (b=t)
  $.
  Taking for example $t=b$, we deduce $\modal (a=b) = \modal(a=a)$,
  and the latter is inhabited by $\eta_{a=a} 1$.
\item Suppose now $p : \modal(a=b)$. Let $\iota$ be the first
  projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
  an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
  (\mu_T b)$, \ie $\prod_{t:T}\modal (a=t) = \modal (b=t)$. The latter
  remains true by univalence.
\end{itemize}
The fact that these two form an equivalence is technical, we refer to
the formalization for an explicit proof.
\end{proof}

Now, let $Q$ be a separated $\Type_{n+1}$, and $f:T \to Q$. Then the
following diagram commutes
$$\xymatrix @C=4em{ 
    \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
    \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
    \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
    & Q &
  }$$
But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
colimit of the closed diagonals diagram, thus there is an universal
arrow $\separated T \to Q$.
%
This is enough to state the following proposition.
\begin{prop}
  $(\separated,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
\end{prop}

To show that $\separated$ is a modality, it remains to show point {\it (v)}. Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be
a sheaf family. We want to show that $\sum_{x:A} (Bx)$ is
separated. Let $E$ be a type, and $\sum_{e:E} (\chi\,e)$ a dense
subobject of E.

@@

$$\xymatrix@C=8em@R=4em{
  \sum_{y:\sum_{e:E} (\chi\,
  e)} x=y_1 \ar@<2pt>[r]^{\qquad \pi_2\circ f\circ\pi_1\circ \pi_1}
\ar@<-2pt>[r]_{\qquad \pi_2\circ g\circ \pi_1\circ \pi_1}
\ar[d]_{\mathrm{dense}}& B\,x \\
  \sum_{y:E} x = y \ar@<2pt>[ur]^{\pi_2\circ f\circ \pi_1} \ar@<-2pt>[ur]_{\pi_2\circ g\circ \pi_1}&
}$$
Using the sepatation property of $B\,x$, one can show that second
components, transported correctly along the first components equality,
are equal. The complete proof can be found in the formalisation.
This proves the following proposition
\begin{prop}
  $(\separated,\mu)$ defines a modality on $\Type_{n+1}$.
\end{prop}

As this modality is just a step in the construction, we do not show
that it is left exact, we will only do it for the sheafification modality.

@@

As any separated type $T$ embeds in $\left(\Type_n^\modal\right)^T$,
it suffices to take the closure of $T$ to get a sheaf, and the unit
$\nu_T:T \to \star T$ is obvious.
\begin{prop}
  $(\star,\nu)$ defines a reflective subuniverse.
\end{prop}
\begin{proof}
  Let $T,Q:\Type_{n+1}$ such that $Q$ is a sheaf. Let $f:T\to Q$.
  Because $Q$ is a sheaf, it is in particular separated ; thus we can
  extend $f$ to $\separated f:\separated T\to Q$.

  But as $\star T$ is the closure of $\separated T$, $\separated T$ is dense
  into $\star T$, so the sheaf property of $Q$ allows to extend
  $\separated f$ to $\star f:\star T \to Q$.

@@

At total, with other files containing technical lemmas, the project
contains 6322 lines.

\nt{describe how many line for each, be clear about what has not been formalized}
